### 2020-10-22 Analysis/Code/Sorokin_JMP_2020.Rmd
Value proposition added

 abstract: |
  \onehalfspacing \noindent 
  I document that firms in online marketplaces use price promotions to facilitate transitions to better review tiers (similar to the number of stars on Amazon.com). I find that firms close to going one tier up are 4-9% more likely to discount. I theorize that two effects could be at play. First, a selection effect arises because customers who buy during a discount could be different from the regular ones, and could potentially leave more positive reviews. Second, a variance effect reflects the idea that positive reviews could help the firm move up a review tier, while negative reviews would keep the tier unchanged, minimizing the downside risk of giving a discount. To test my hypotheses, I estimate a simple structural model of demand and reviewing behavior, and develop a new approach to estimating demand from data on product usage. I do not find evidence that the selection effect is positive. I find evidence of the importance of the variance effect: when the selection effect is controlled for, firms close to downgrading their review tier are 6% less likely to give a discount, consistent with preferring less variance. I also find that consumers are significantly more likely to leave reviews during a discount. Additional findings include estimates of the causal effect of reviews on sales and equilibrium discount elasticities in an important market not previously studied. \thispagestyle{empty}

@@

In his book “Economics for the Common Good” Jean Tirole proclaims the economy we live in to be governed by “Economics of Attention” [-@Tirole17, p.379]. A simple search for wine glasses on Amazon.com results in more than 5,000 options, leading to a choice problem that is virtually impossible to solve satisfyingly without guidance. Such guidance is now routinely provided by different types of recommendation systems. Many marketplaces or third party websites allow consumers to review products and services that they purchased, and then aggregate these reviews into a rating, greatly facilitating the comparison between alternatives. These ratings are often given a simple graphical representation; for example, a number of stars between 0 and 5 is a common way to split the alternatives into *review tiers*. Such review tiers are used to review household items (Amazon), movies (IMDb), restaurants (Yelp), doctors (RateMDs), and many other products and services. Both an extensive literature in marketing, and anecdotal evidence from these markets, suggest that good reviews are important, if not crucial, for the success of the business. Not surprisingly, firms take an active approach to managing their reviews. Firms solicit reviews from their clients, publicly respond to reviews, and even leave fake reviews for themselves or their competitors. In this paper I study a novel form of review management. I show that firms can use discounting in order to facilitate *transitions* between review tiers. Imagine a product that is several good reviews short of obtaining an additional star. I ask: would it be profitable for such a product to run a price promotion?

With a help of a stylized model, I identify two effects that influence this decision: the *selection* effect and the *variance* effect. Customers who purchase the good during a discount are likely to be different from the ones who pay the full price, and could potentially leave different reviews. If such reviews tend to be more favorable, then the induced positive selection effect creates incentives for the product to go on sale, as new customers are likely to leave positive reviews and help the product advance to the next review tier. However, these new customers could also prove to be a worse match with the product and end up leaving less favorable reviews. Thus, the sign of the selection effect is ultimately an empirical question. The variance effect reflects the idea that a product that is close to a transition is willing to gamble on upgrading its review tier. Additional positive reviews catapult it to the next review bin, which promises higher sales, while a few negative reviews might not affect its review status---and, therefore, sales---by much. While extra variance is good for products that are on the verge of a tier upgrade, it could be harmful for products that are close to sliding one review tier down. Fearing to lose that precious star, such products would prefer to delay running a price promotion, unless the selection effect is positive and strong.

To quantify all these effects, I analyze data from Steam, a major marketplace for computer video games. I build and estimate a tractable structural model of demand and reviewing behavior on Steam. The demand model lays the groundwork for my analysis, as it helps me confirm the importance of better review tiers for sales on Steam. I find a premium of as high as 10% of being in the best bin and a penalty of up to 13% for not having reviews, compared to a mediocre review tier. My estimates also highlight how impactful discounts are in this market, with an average discount being accompanied by a 25% spike in sales. Thus, firms on Steam have both the ability to use discounts to attract new users, and the incentives to upgrade their review tiers, given the differential contribution of various tiers to sales.

@@

>>>>

 abstract: |
  \onehalfspacing \noindent 
  This paper documents that firms in online marketplaces use price promotions to facilitate transitions to better review tiers (similar to the number of stars on Amazon.com). I find that firms close to going one tier up are 4-9% more likely to discount. Two effects could be at play. First, a selection effect arises because customers who buy during a discount could be different from the regular ones, and could potentially leave more positive reviews. Second, a variance effect reflects the idea that positive reviews could help the firm move up a review tier, while negative reviews would keep the tier unchanged, minimizing the downside risk of giving a discount. In order to understand the relative contributions of these effects, I estimate a tractable structural model of demand and reviewing behavior. I find evidence of the importance of the variance effect: when the selection effect is controlled for, firms close to downgrading their review tier are 6% less likely to give a discount, consistent with preferring less variance. I also find that consumers are significantly more likely to leave reviews during a discount. The average selection effect turns out to be small. Additional findings include developing a new approach to estimating demand from data on product usage, estimates of the causal effect of reviews on sales and equilibrium discount elasticities in an important market not previously studied. 
  \thispagestyle{empty}

@@

In his book “Economics for the Common Good” Jean Tirole proclaims the economy we live in to be governed by “Economics of Attention” [-@Tirole17, p.379]. A simple search for wine glasses on Amazon.com results in more than 5,000 options, leading to a choice problem that is virtually impossible to solve satisfyingly without guidance. Such guidance is now routinely provided by different types of recommendation systems. Many marketplaces or third party websites allow consumers to review products and services that they purchased, and then aggregate these reviews into a rating, greatly facilitating the comparison between alternatives. These ratings are often given a simple graphical representation; for example, a number of stars between 1 and 5 is a common way to split the alternatives into *review tiers*. Such review tiers are used to review household items (Amazon), movies (IMDb), restaurants (Yelp), doctors (RateMDs), and many other products and services. Both an extensive literature in marketing, and anecdotal evidence from these markets, suggest that good reviews are important, if not crucial, for the success of the business. Not surprisingly, firms take an active approach to managing their reviews. Firms solicit reviews from their clients, publicly respond to reviews, and even leave fake reviews for themselves or their competitors. In this paper I study a novel form of review management: I show that firms can use discounting in order to facilitate *transitions* between review tiers. Imagine a product that is several good reviews short of obtaining an additional star. My question is: would it be profitable for such a product to run a price promotion?

With a help of a stylized model, I identify two effects that influence this decision: the *selection* effect and the *variance* effect. Customers who purchase the good during a discount are likely to be different from those who pay the full price, and could potentially leave different reviews. If such reviews tend to be more favorable, then the induced positive selection effect creates incentives for the product to go on sale, as new customers are likely to leave positive reviews and help the product advance to the next review tier. However, these new customers could also prove to be a worse match with the product and end up leaving less favorable reviews. Thus, the sign of the selection effect is ultimately an empirical question. The variance effect reflects the idea that a product that is close to a transition is willing to gamble on upgrading its review tier. Additional positive reviews catapult it to the next review bin, which promises higher sales, while a few negative reviews might not affect its review status---and, therefore, sales---by much. While extra variance is good for products that are on the verge of a tier upgrade, it could be harmful for products that are close to sliding one review tier down. Fearing to lose that precious star, such products would prefer to delay running a price promotion, unless the selection effect is positive and strong.

To quantify all these effects, I analyze data from Steam, a major marketplace for computer video games. I build and estimate a tractable structural model of demand and reviewing behavior on Steam. The demand model lays the groundwork for my analysis, as it helps me confirm the importance of better review tiers for sales on Steam. I find a premium of as high as 10% of being in the best bin and a penalty of up to 13% for not having reviews, compared to a mediocre review tier. My estimates also highlight how impactful discounts are in this market, with an average discount being accompanied by a 38% spike in sales. Thus, firms on Steam have both the ability to use discounts to attract new users, and the incentives to upgrade their review tiers, given the differential contribution of various tiers to sales.

@@

This paper contributes to the ongoing research effort to improve the design of  recommendation systems. Theoretical advancements have been recently made in understanding  separate components of the ideal system, such as exploration of new products vs. exploitation of the established options (@cheHorner18), creation of incentives for firms to invest in quality (@AperjisJohari10), promoting competition and reducing barriers to entry (@Vellodi20), to name just a few topics and contributions. At the heart of every such inquiry lies a particular strategic interaction between firms, users, and the platform. My paper provides empirical evidence on what some of these interactions result in. I show that firms respond to the tier structure of review systems by changing their pricing behavior around the thresholds, and that they use pricing to affect their review process. My estimates can inform future work that will synthesize theoretical and data-driven insights in order to make concrete design recommendations. For example, @Vellodi20 shows that barriers to entry resulting from the low speed of review accumulation by new firms calls for a review system that censors the reviews of the most established products; however, when firms can offer lower prices to build up the crucial review base, the need for the design intervention is diminished. My paper directly informs this question by quantifying two fundamental forces involved: the frequency with which firms change prices in response to their review status, and the response of the reviewing behavior to these changes in prices.

## Literature Review
